---
title: 'Assignment: Course Project 1'
author: "Ashish"
date: "10 February 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Analysis of Personal Movement Using Activity Monitoring Devices

## Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This project makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day. The goal of this project is to write a report that answers the questions detailed below.

##Loading and preprocessing the data

####1Reading the data from the "activity.csv"

```{r}
data <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")
```
####2 - Looking at a summary for the dataset using "summary" and "str" methods:

```{r}
summary(data)
```

### Looking at the first 6 rows of the dataset
```{r}
head(data)
```

### Converting the "date" variable to a Date classe and the "interval" variable to a factor:
```{r}
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$interval <- factor(data$interval)
```
 
##What is mean total number of steps taken per day?
```{r}
NA_index <- is.na(as.character(data$steps))
data_no_NA <- data[!NA_index,]
head(data_no_NA)
```

### Aggregating the number of steps taken each day:
```{r}
steps_each_day <- aggregate(steps ~ date, data = data_no_NA, sum)

colnames(steps_each_day) <- c("date", "steps")
```

###Making a histogram of the total number of steps taken each day:
```{r}
hist(as.numeric(steps_each_day$steps), breaks = 20, col = "red", xlab = "Number of Steps", 
   main= "Histogram  of the total number of steps taken each day")

mean(steps_each_day$steps)
median(steps_each_day$steps)
```
###What is the average daily activity pattern?
```{r}
hist(as.numeric(steps_each_day$steps), breaks = 20, col = "red", xlab = "Number of Steps", 
     main= "Histogram of the total number of steps taken each day")

steps_per_interval <- aggregate(data_no_NA$steps, by=list(interval=data_no_NA$interval), FUN=mean)
colnames(steps_per_interval) <- c("interval", "average_steps")
plot(as.integer(levels(steps_per_interval$interval)), steps_per_interval$average_steps, type="l",
     xlab = "Interval", ylab = "Average Number of Steps", main = "Average Daily Activity Pattern", 
    col ="blue")
```
lobrary
### 5-minute interval that contains the maximum number of steps:
```{r}
max_steps <- max(steps_per_interval$average_steps)
max_steps
intervale_max_steps<-steps_per_interval[which.max(steps_per_interval$average_steps),]$interval
intervale_max_steps
```

###Imputing missing values
```{r}
sum(is.na(as.character(data$steps)))
sum(is.na(as.character(data$date)))
sum(is.na(as.character(data$interval)))
```

###The strategy for filling in all of the missing values in the dataset. Missing values are replaced by the mean of that 5-minute interval.

```{r}
NA_index <- which(is.na(as.character(data$steps)))
complete_data <- data
complete_data[NA_index, ]$steps<-unlist(lapply(NA_index, FUN=function(NA_index){
                steps_per_interval[data[NA_index,]$interval==steps_per_interval$interval,]$average_steps
                }))
```

###Creating a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
summary(complete_data)
str(complete_data)

steps_each_day_complete <- aggregate(steps ~ date, data = complete_data, sum)

colnames(steps_each_day_complete) <- c("date", "steps")

hist(as.numeric(steps_each_day_complete$steps), breaks = 20, col = "red", xlab = "Number of Steps", main= "Histogram of the total number of steps taken each day")


```

###Calculating the mean and median total number of steps taken per day for the complete dataset:
```{r}
mean(steps_each_day_complete$steps)
median(steps_each_day_complete$steps)

#Creating a factor variable "day "to store the day of the week:
complete_data$day <- as.factor(weekdays(complete_data$date))

#Creating a logical variable "is_weekday" (weekday=TRUE, weekend = FALE) :
complete_data$is_weekday <- ifelse(!(complete_data$day %in% c("Saturday","Sunday")), TRUE, FALSE) 


#Calculating the average number of steps for weekdays
weekdays_data <- complete_data[complete_data$is_weekday,]
steps_per_interval_weekdays <- aggregate(weekdays_data$steps, by=list(interval=weekdays_data$interval), FUN=mean)


#Calculating the average number of steps for weekends
weekends_data <- complete_data[!complete_data$is_weekday,]
steps_per_interval_weekends <- aggregate(weekends_data$steps, by=list(interval=weekends_data$interval), FUN=mean)

#Adding columns names
colnames(steps_per_interval_weekdays) <- c("interval", "average_steps")
colnames(steps_per_interval_weekends) <- c("interval", "average_steps")
#Adding a column to indecate the day
steps_per_interval_weekdays$day <- "Weekday"
steps_per_interval_weekends$day <- "Weekend"

#Merging the two togather
week_data <- rbind(steps_per_interval_weekends, steps_per_interval_weekdays)
#Converting the day variabke to a factor
week_data$day <- as.factor(week_data$day)

#Making the plot
library(lattice)
xyplot(average_steps ~  interval | day, data = week_data, layout = c(1,2), type ="l", ylab="Number of Steps")


```
